---
title: "Feature Selection of Combined Datasets"
author: "Craig McLean"
date: "5/3/2020"
output: html_document
---

Observation:

From the procrustes analysis, there appears to be three clusters related to 
distinct wells. 

Plan:

1) I'll run an NMF on both datasets
2) I'll select features from bins first (look for organisms driving change)
3) I'll look for pathways that make these organisms distinct
4) I'll look for evidence of each pathway within metabolites NMF profiles from
each well

# \Filtering Genomics Data

```{r loading data}
bins <- read.csv(here::here("data/2020-05-03/bin_rpkm_filtered.csv"))
taxa <- read.csv(here::here("data/2020-05-03/bin_taxons_filtered.csv"))
```

```{r}
library(NMF)
binNMF <- nmf(x = as.matrix(bins[,-1]), rank = 3)
  
keyFeatures <- extractFeatures(binNMF, method = 0.99)
linkedFactors <- apply(coef(binNMF), 1, function(row) {
    row > 0.01
})
print(linkedFactors)
rm(linkedFactors, binNMF)
```

distribution of samples across rank space of NMF
2) NSHQ14 and WAB71
1) WAB55
3) WAB105 and WAB 104

The order is not set in stone

```{r Learning key taxons}
nshq <- bins[keyFeatures[[1]],1]
wab55 <- bins[keyFeatures[[2]],1]
wab105 <- bins[keyFeatures[[3]],1]

taxa$keyWell <- ""
taxa$keyWell[taxa$Bin.Id %in% gsub("-", "_", nshq)] <- "nshq"
taxa$keyWell[taxa$Bin.Id %in% gsub("-", "_", wab55)] <- "wab55"
taxa$keyWell[taxa$Bin.Id %in% gsub("-", "_", wab105)] <- "wab105"
taxa <- taxa[taxa$keyWell != "",]

write.csv(taxa, file = here::here("data/2020-05-03/key_well_taxa.csv"),
          row.names = F)
rm(wab105,wab55, nshq, keyFeatures, bins)
```

# Pathways for key microbes

```{r loading Gene data}
binKos <- read.csv(here::here("data/2020-04-12/Oman_bins_kos.csv"),
                   stringsAsFactors = F)

# filtering data to only consider key bins
binKos <- binKos[,c(1:5,which(colnames(binKos) %in% taxa$Bin.Id))]
binKos <- binKos[!is.na(binKos$Metabolism),]
binKos <- binKos[!duplicated(binKos$KO_ID),]
binKos[is.na(binKos)] <- 0

count <- t(binKos[,-c(1:5)])
colnames(count) <- binKos$KO_ID
count[count > 1] <- 1
count <- as.data.frame(count)
count$well <- taxa$keyWell
#count$well[count$well != "nshq"] <- "other"
count$well <- as.factor(count$well)

```


I think a random forrest could be used to identify genes that seperate the types of wells

```{r}
library(randomForest)
require(caTools)


pc <- princomp(t(count[,-ncol(count)]))

plot(pc$scores[,5],pc$scores[,4])

rf <- randomForest(
  
    # denotes what I'm actually predicting for
    well ~ .,
    data=count,
    importance = T
)

```

